Expected Dependency Pair Match:\\ Predicting translation quality with expected syntactic structure

### 2009-05-15 mtjournal.tex
changes to include tuning plot in text. much text written.

We train a model using features from one or more of the following
sets:
\begin{description}
\item[E]: the fully syntactic features from the DPM family.
  Specifically, this feature set includes precision and recall scores
  for the $dl$, $lh$, $dlh$, and $dh$ decompositions of the syntactic
  structure, using $n=50$ and $\gamma=0.25$.  The \textbf{E} set
  includes 8 features.
\item[N]: the non-syntactic features from the DPM family;
  specifically, precision and recall for $1g$ and $2g$ decompositions.
  The \textbf{N} set includes 4 features.
\item[T]: the features from basic TERp, excluding paraphrases, which include a separate cost
  for each of matches, insertions, deletions, substitutions, shifts, and
  synonym matches and stem matches. The \textbf{T} set includes 7 features.
\item[P]: features from the automatically-derived TERp paraphrase
  table. These are four features which are a function of of the edit
  distance between the two phrases in each paraphrase and the
  probability of the paraphrase.
\item[B]: two ``length penalties'' one is assigned the count of
  hypothesis words beyond the count of reference words, and one
  assigned the count of reference words beyond the count of hypothesis
  wordsbrevity/prolixity (2 features: one for longer-than-ref, one for
  shorter-than-ref).
  % MO: out of curiosity, why do you need 2 features?
  We assign two features here to let the system separately penalize
  ``too short'' translations and ``too long''.
\end{description}

@@

% tuning target: weighted correlation with mean-removed segment HTER

% [we use weighted correlation to avoid emphasis on short segments
% (which hurts document-level and system-level utility) ]
In figure~\ref{fig:tuneresults}, we plot the Pearson's $r$ (with 95\%
confidence interval) for the results on the two test sets (each
trained on one and tested on the other).
\begin{figure}
  \centerline{{one figure}\qquad{}{two figure}}
  \caption{Figure with two tune-set plots will go here.}
  \label{fig:tuneresults}
\end{figure}

% MO suggestions:

%% do n-fold with >2 N -- better robustness to variation?

%% look at variation in weights on 2-fold case   
\begin{verbatim}
   == Tuned For Weighted Mean Removed Pearson
     - Examining WGT_MEAN_RM_SEG_PEAR ==
            Average
      Cond   Test  
     ---------------
      ETPB   0.4096
      ETP    0.4038
      TPB    0.4010
      TP     0.3946  > 0.416
      TPNB   0.3940
      TPN    0.3891
      ETB    0.3871
      ET     0.3827  > 0.404
      TB     0.3815  > 0.403
      ETNB   0.3814
      ETN    0.3804
      T      0.3758  > 0.398
      TNB    0.3708
      TN     0.3680  > 0.390
      TER    0.353   > 0.375
      E      0.3287  > 0.352
      EB     0.3267
      ENB    0.3149
      NB     0.3058  > 0.329
      EN     0.2973
      N      0.2636
      BLEU   0.244
      B      0.1638
\end{verbatim}
(maybe don't report all these numbers!)

TO DO: discuss this as a trend, despite lack of significance? plot charts?

point: E+T $>$ T, E+TB $>$ TB, E+TP $>$ TP; information is available
in syntax that is not captured in the other measures.

point: syntax is *not* just an expensive way to get at n-grams;
N-grams are not the same -- TP $>$ TP+N, but TP $<$ TP+E

question: this is kind of like the $\mu_{PR}[\cdot]$ --- seems like it
also doesn't work so well. compare to EDPM as-is

talk about why EN is $<$ EDPM (is it?)

>>>>

We train a model using features from one or more of the following
sets:
\begin{description}
\item[E]: the fully syntactic features from the DPM family.
  Specifically, this feature set includes error-counts
  ($\frac{|\textrm{hyp}| - |\textrm{correct}|}{|\textrm{ref}|}$ and
  $\frac{|\textrm{ref}| - |\textrm{correct}|}{|\textrm{ref}|}$) 
  for the $dl$, $lh$, $dlh$, and $dh$ decompositions of the syntactic
  structure, using $n=50$ and $\gamma=0.25$.  The \textbf{E} set
  includes 8 features.
\item[N]: the non-syntactic features from the DPM family;
  specifically, error-counts for the $1g$ and $2g$ decompositions.
  The \textbf{N} set includes 4 features.
\item[T]: the features from basic TERp, excluding paraphrases, which include a separate cost
  for each of matches, insertions, deletions, substitutions, shifts, and
  synonym matches and stem matches. The \textbf{T} set includes 7 features.
\item[P]: features from the automatically-derived TERp paraphrase
  table. These are four features which are a function of of the edit
  distance between the two phrases in each paraphrase and the
  probability of the paraphrase.
\item[B]: two ``length penalties'' one is assigned the count of
  hypothesis words beyond the count of reference words, and one
  assigned the count of reference words beyond the count of hypothesis
  wordsbrevity/prolixity (2 features: one for longer-than-ref, one for
  shorter-than-ref).
  % MO: out of curiosity, why do you need 2 features?
  We assign two features here to let the system separately penalize
  ``too short'' translations and ``too long''.
\end{description}

@@

% tuning target: weighted correlation with mean-removed segment HTER
\subsection{Effects of features on tuning}
% [we use weighted correlation to avoid emphasis on short segments
% (which hurts document-level and system-level utility) ]
In figure~\ref{fig:tuneresults}, we plot the Pearson's $r$ (with 95\%
confidence interval) for the results on the two test sets (each
trained on one and tested on the other).
\begin{figure}
  \begin{center}
    \includegraphics[scale=0.45]{tuning.eps}
  \end{center}
  \caption{Pearson's $r$ values for the two halves of the training
    set, with 95\% confidence intervals. }
  \label{fig:tuneresults}
\end{figure}
%
For each set of features, we plot the $r$ (and confidence) on both
halves of the corpus (for each half, tuned on the other half). In the
leftmost group of the figure, we include baselines: EDPM and BLEU and
TER.  Next, the tuned E feature, which (by itself) does not perform as
well as TER or EDPM.  
%
The remaining groups of three show that for each set of features, as E
(and EB) are included with others, correlation improves (though not
significantly).  These all move in the same direction. By the same
measurement, P features help beyond T -- as reported elsewhere -- but
N features do not seem to find a better correlation than TP features.
% MO suggestions:

Overall, we find that this approach for combining the underlying idea
of EDPM with the underlying ideas of TERp is promising, in that the E
features offer small (though non-significant) improvements to the
features already within TERp. However, this combination approach still
has more in common with the $\mu_{PR}$ combination method (explored in
section~\ref{sec:faexpts}) than with EDPM, which suggests a reason
that the EN features --- though tuned --- do not outperform EDPM.
%% do n-fold with >2 N -- better robustness to variation?
